---
title: "Tidyverse Problem Set"
author: "Dae Hyun Lee"
date: "October 4, 2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
urlcolor: blue
---

```{r setup, include=FALSE}
library(knitr)
  options(tinytex.verbose = TRUE)
  opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidyr)
library(ggplot2)
library(kableExtra)
library(magrittr)
library(esquisse)
library(gridExtra)
library(shiny)
library(dplyr)
library(AER)
```

### Problem 1

Load the gapminder data from the gapminder package.
```{r echo=FALSE}
library(gapminder)
data(gapminder)
```

How many continents are included in the data set?
```{r echo=FALSE}
levels(gapminder$continent)
```
_There are total 5 continents in the data set._

How many countrys are included?  How many countries per continent?
```{r echo=FALSE}
head(fct_unique(gapminder$country))
```
_There are total 142 courtrys are included in the data set._

Using the gapminder data, produce a report showing the continents in the dataset, total population per continent, and GDP per capita.  Be sure that the table is properly labeled and suitable for inclusion in a printed report.
```{r echo=FALSE}
#year <- gapminder$year
#unique(year) ## check what years are in the data, and decide to narrow the report to 1992.
gapminder1992 <- filter(gapminder, year == 1992)  ##filter data to year in 1992
gap1992 <- mutate(gapminder1992, gdp = pop * gdpPercap)       ##create a variable containing GDP
totalp <- aggregate(gap1992[c("pop")], list(Continent=gap1992$continent), sum)   ##population by continents in 1992
totalg <- aggregate(gap1992[c("gdp")], list(gap1992$continent), sum)          ##overall GDP by continents in 1992
gpercapita <- totalg$gdp / totalp$pop                        ##overall GDP / overall population by continents
totalGP <- mutate(totalp, fixedgdpPercap = gpercapita)     ##Combine two variables together in data
d1 <- as_tibble(totalGP)
kable(d1, digits = 2, align = 'c',
      col.names = c("Continent", "Total Population", "GDP per Capita"), caption = "Total population and per capita GDP of each continent in 2002") %>% 
  kable_styling()

## Why does this table keep appears to show in the next page in such abnormal format
```

Produce a well-labeled table that summarizes GDP per capita for the countries in each continent,  contrasting the years 1952 and 2007.
```{r echo=FALSE}
## unique(gapminder$continent)
gapminder1 <- mutate(gapminder, gdp = gapminder$gdpPercap * gapminder$pop)
gap1952 <- filter(gapminder1, year == 1952)
gap2007 <- filter(gapminder1, year == 2007)
get1 <- function(data1){
  get.gdp <- function(data0){
    sum(data0$gdp)/sum(data0$pop)
  }
  x1 <- filter(data1, continent == "Africa")
  x2 <- filter(data1, continent == "Europe")
  x3 <- filter(data1, continent == "Asia")
  x4 <- filter(data1, continent == "Americas")
  x5 <- filter(data1, continent == "Oceania")
  return(c(get.gdp(x1), get.gdp(x2), get.gdp(x3), get.gdp(x4), get.gdp(x5)))
}
tb.gdp2007 <- get1(gap2007)
tb.gdp1952 <- get1(gap1952)
cont <- c("Africa", "Europe", "Asia", "Americas", "Oceania")
ex1.table2 <- tibble(
  `continent` = cont,
  `gdp1952` = tb.gdp1952,
  `gdp2007` = tb.gdp2007
)
kable(ex1.table2, digits = 2,
      col.names = c("Continent", "GDP per capita in 1952", "GDP per capita in 2007"), 
      caption = "GDP per capita by continents in 1952 and in 2007", align = 'c') %>%
  kable_styling(latex_options = 'hold_position',font_size = 12,full_width = F,position = "center") %>%
  column_spec(1,bold = T)
```

Product a plot that summarizes the same data as the table. There should be two plots per continent.
_wdwd_

Which countries in the dataset have had periods of negative population growth?
_oo_

Illustrate your answer with a table or plot.
_oo_

Which countries in the dataset have had the highest rate of growth in per capita GDP?
_k_

Illustrate your answer with a table or plot.
_j_


\newpage

### Problem 2

The data for Problem 2 is the Fertility data in the AER package.  This data is from the 1980 US Census and is comprised of date on married women aged 21-35 with two or more children.  The data report the gender of each woman's first and second child, the woman's race, age, number of weeks worked in 1979, and whether the woman had more than two children.

```{r echo=FALSE}
data("Fertility")
Fertility <- as_tibble(Fertility)
```

There are four possible gender combinations for the first two Children.  Product a plot the contracts the frequency of these four combinations. Are the frequencies different for women in their 20s and wemen who are older than 29?
```{r echo=FALSE}
freq <- Fertility %>% unite("combined", gender1, gender2, sep="&")
freq %<>% mutate(age.2030=ifelse(age>29, "age_30s", "age_20s"))
ggplot(freq) +
  aes(x=combined, fill=age.2030) +
  geom_bar(position=position_dodge(width=1)) +
  labs(x="Gender of children", y="Frequency", title="Frequency of combinations", fill="Mom's age") +
  theme_classic()
```

_It is clear that there are differences in frequency between two age groups._

Produce a plot that contrasts the frequency of having more than two children by race and ethnicity. 
```{r echo=FALSE}
## Filter women with more than two children
fert1 <- filter(Fertility, morekids=="yes")
## Filter women by race
race.af <- filter(fert1, afam=="yes")
race.af$race <- "Afam"
race.his <- filter(fert1, afam=="no" & hispanic=="yes")
race.his$race <- "Hisp"
race.non <- filter(fert1, afam=="no" & hispanic=="no" & other=="yes")
race.non$race <- "Other"
race.cau <- filter(fert1, afam=="no" & hispanic=="no" & other=="no")
race.cau$race <- "Cauc"
## Combine data together
fert2 <- bind_rows(race.af, race.his, race.non, race.cau)
## Add column whose age is whether above 29 or below 30.
fert2 %<>% mutate(age.2030=ifelse(age>29, "age_30s", "age_20s"))
## Plot
ggplot(fert2) +
  aes(x=race, fill=age.2030) +
  geom_bar() +
  labs(x="Race", y="Frequency", title="Frequency of having more than two children by race", fill="Mom's age") +
  theme_classic()
```

### Problem 3

Use the mtcars and mpg datasets.~~
How many times does the letter "e" occur in mtcars rownames?
```{r}
data(mtcars)
data(mpg)
```

How many cars in mtcars have the brand Merc?

How many cars in mpg have the brand("manufacturer" in mpg) Merc?

Contrast the mileage data for Merc cars as reported in mtcars and mpg.  Use tables, plots, and a short explaination.


### Problem 4

Install the babynames package.
```{r}
library(babynames)
data(babynames)
babyn <- as_tibble(babynames)
babyn
```
Draw a sample of 500,000 rows from the babynames data
```{r}
##hist(babyn$year)
s <- sample(x = 1:1924665, size = 500000, replace = FALSE)
babyn_500000 <- babyn %>% slice(s)
babyn_500000
```
Produce a tabble that displays the five most popular boy names and girl names
in the years 1880,1920, 1960, 2000.

What names overlap boys and girls?

What names were used in the 19th century but have not been used in the 21sth century?

Produce a chart that shows the relative frequency of the names "Donald", "Hilary", "Hillary", "Joe", "Barrack",  over the years 1880 through 2017.






