---
title: "Slides for the Mid-term Assignment"
author: "Dae Hyun Lee"
date: "10/21/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)                ## loading the entire knitr package
library(ggplot2)              ## ggplot2 is the only tidyverse library that I need
library(esquisse)             ## use esquisse to setup the basic plots
library(WDI)
library(kableExtra)
library(magrittr)
library(gridExtra)
library(AER)
library(corrplot)
library(cluster)
library(rattle)
library(factoextra)
library(FactoMineR)
load("/Users/daylelee/Desktop/MA 615 - R/MA615__Git_Repository/-Midterm Project/MA615_midterm/wdi_dat_pov.RData")
```

## WDI (World Development Indicators) - World Bank
Search, extract and format data from the World Bank's World Development Indicators.

```{r}
## unique(wdi_dat_pov$country)  # show countries in the data
names(wdi_dat_pov)  #show variables in the data
```

```{r echo=FALSE, include=FALSE}
# cleaning data for factor and cluster analysis 
data <- wdi_dat_pov[,c(3:9)] # data only with numeric values
data00 <- na.omit(wdi_dat_pov[,c(2:9)]) # data without any NA values
data0 <- na.omit(data) # data without any NA values, but contains country names
```

# Exploratory Factor Analysis

## Correlation Plot

```{r echo=FALSE}
scaled_dat <- scale(data0)
corr_dat <- cor(scaled_dat)
colnames(corr_dat) <- c("Year", "GDP", "GDP per capita", "Population","Infant Mortality", "Life Expectancy", 
                         "Maternal Mortality")

rownames(corr_dat) <- c("Year", "GDP", "GDP per capita", "Population","Infant Mortality", "Life Expectancy", 
                         "Maternal Mortality")

corrplot(corr_dat, method = "color")
```

## Scree plot of eigenvalues after factor

```{r echo = FALSE}
screeplot <- function(data0, nc = 7, seed = 42){
    sc <- nrow((data0)-1)*sum(apply(data0, 2, var))
    for (i in 2:nc){
      set.seed(seed)
      sc[i] <- sum(kmeans(data0, centers = i)$withinss)}
    plot(1:nc, sc, type = "b", xlab = "Number of Clusters", ylab = "Eigenvalue")}
screeplot(scaled_dat, nc = 7)
```

# Cluster Analysis

## Clustering Figure 1
Grouping a set of data objects into clusters

```{r, echo = FALSE}
scdat <- kmeans(scaled_dat, 3)
clusplot(scaled_dat, scdat$cluster, main = '2D plot of Clusters: 3 clusters', color = TRUE, shade = TRUE, labels = 2, lines = 0)
```

## Clustering Figure 2
Determining the optimal number of clusters

```{r, echo = FALSE}
fviz_nbclust(scaled_dat, kmeans, method = "gap_stat")
```

## Clustering Figure 3
Hierarchical clustering through euclidean distance

```{r echo = FALSE}
scdat2 <- dist(scaled_dat, method = "euclidean")
fit_1 <- hclust(scdat2, method = "ward.D")
plot(fit_1)
groups <- cutree(fit_1, k = 3)
rect.hclust(fit_1, k = 3, border = "red")
```

# EDA for Variables

## Negatively correlated variables

```{r, echo = FALSE, message = FALSE, warning = FALSE}
ggplot(data00) +
 aes(x = GDP_perCap, y = infant_mortality) +
 geom_point(aes(color = country), size = 1.8, colour = "#ef562d") +
 geom_smooth(geom="line", alpha=0.4, size=0.9, span = 0.75) +
 scale_x_continuous(trans = "log") +
 scale_y_continuous(trans = "log") +
 labs(x = "GDP per Capita", y = "Infant Mortality", title = "Negatively correlated variable pair") +
 theme_classic()
```

## Positively correlated variables

```{r, echo = FALSE, message = FALSE, warning = FALSE}
ggplot(data00) +
 aes(x = GDP_perCap, y = life_expectancy) +
 geom_point(aes(color = country), size = 1.8, colour = "#ef562d") +
 geom_smooth(geom="line", alpha=0.4, size=0.9, span = 0.75) +
 scale_x_continuous(trans = "log") +
 scale_y_continuous(trans = "log") +
 labs(x = "GDP per Capita", y = "Life Expectancy", title = "Positively correlated variables") +
 theme_classic()
```

# EDA on WDI related to poverty

## Infant Mortality

```{r ,echo=FALSE, message = FALSE, warning = FALSE}
ggplot(subset(wdi_dat_pov, year == 2008), aes(x = GDP_perCap, y = infant_mortality)) + geom_point() +
  labs(x="GDP per Capita", y="Infant Mortality Rate", 
       title="Figure 0: Relationship between Infant Mortality Rate and GDP per Capita in year 2008") +
  theme_classic()
```

## Life Expectancy

```{r ,echo=FALSE}
co <- c("United States", "Rwanda", "Mongolia", "Pakistan", "Lao PDR", "Bhutan", "Malaysia", "Brazil", "Ireland", "Japan", "China","Sweden", "Netherlands","Korea, Rep.")
lifexp_c <- subset(wdi_dat_pov, country %in% co) # subset specific countries from the data
lifexp_c0 <- subset(lifexp_c, year == 2016)

ggplot(lifexp_c0) +
 aes(x = GDP_perCap, y = life_expectancy, colour = country) +
 geom_point(size = 2.56) +
 geom_text(aes(label = country), size=2, nudge_y = 0.8) +
 scale_color_hue() +
 labs(x = "GDP per Capita (USD)", y = "Life Expectancy (year)", title = "Figure 1: Life Expectancy and GDP per Capita by Country", subtitle = "(2016)", color = "Country") +
 scale_x_continuous(limits = c(0, 70000)) +
 theme_classic()
```

## Infant Mortality

```{r ,echo=FALSE}
ggplot(lifexp_c0) +
 aes(x = GDP_perCap, y = infant_mortality, colour = country) +
 geom_point(size = 2.56) +
 geom_text(aes(label = country), size=2, nudge_y = 1.9) +
 scale_color_hue() +
 labs(x = "GDP per Capita (USD)", y = "Infant Mortality Rate per 1,000 births", title = "Figure 2: Infant Mortality Rates by GDP per Capita", subtitle = "(2016)", color = "Country") +
 scale_x_continuous(limits = c(0, 70000)) +
 theme_classic()
```

## Simple Interactivity

```{r ,echo=FALSE, message = FALSE, warning = FALSE}
library(plotly)
g <- ggplot(subset(lifexp_c, year > 1999 & year <= 2015), aes(x = GDP_perCap, y = infant_mortality, color = country == "Korea, Rep.", tooltip = country)) +
  geom_point(size = 0.8) +
  facet_wrap(~ year) + theme(legend.position="none") +
  labs(x="GDP per Capita (USD)", y="Infant Mortality Rate per 1,000 births", title= "Figure 3: GDP per capita vs Infant Mortality Rates by years",
       subtitle = "The colored dot represents Korea, Rep.", color = "Country") +
  scale_color_manual(labels = c("T999", "T888"), values = c("darkblue", "red")) +
  theme_dark()
ggplotly(g)
```