---
title: "Midterm Project"
author: "Dae Hyun Lee"
date: "10/20/2019"
output: pdf_document
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)
library(tidyverse)
library(kableExtra)
library(magrittr)
library(esquisse)
library(gridExtra)
library(shiny)
library(AER)
```

# Get Data  
read WDI data
```{r ,echo=FALSE}
library(WDI)
new_wdi_cache <- WDIcache()
```

Next, search for the data that I am interested in.
Try to find out the IDs for the indicators for GDP, GDP growth, per capita GDP, life expectancy, infant mortality and etc,.  
WDI search
```{r ,echo=FALSE}
# WDIsearch("gdp.", cache = new_wdi_cache)
# WDIsearch("gdp.*capita.*US\\$", cache = new_wdi_cache)
# WDIsearch("^mortality.*rate.*infant", cache = new_wdi_cache)
# WDIsearch("life expectancy at birth.*total", cache = new_wdi_cache)
# WDIsearch("maternal mortality")
```

```
These are the list of indicators that I found for EDA
"NY.GDP.MKTP.CD"             "GDP (current US$)" 
"NY.GDP.MKTP.KD.ZG"          "GDP growth (annual %)" 
"NY.GDP.PCAP.CD"             "GDP per capita (current US$)"
"NY.GDP.PCAP.KD.ZG"          "GDP per capita growth (annual %)"

"SP.DYN.IMRT.IN"             "Mortality rate, infant (per 1,000 live births)"
"SP.DYN.LE00.IN"             "Life expectancy at birth, total (years)"
"SH.STA.MMRT.NE"             "Maternal mortality ratio (national estimate, per 100,000 live births)"

Maternal mortality ratio is the number of women who die from pregnancy related causes while pregnant or within 42 days of pregnancy termination per 100,000 live births."

"SP.POP.TOTL"                "Population, total"
"SP.URB.TOTL.IN.ZS"          "Urban population (% of total population)"
"SP.RUR.TOTL.ZS"             "Rural population (% of total population)"

"SH.XPD.CHEX.GD.ZS"          "Current health expenditure (% of GDP)"            data only available from 2000 ~ 2016
"1.1_ACCESS.ELECTRICITY.TOT" "Access to electricity (% of total population)"    few missing data
```


# Clean up the data
Extract the data that I am interested in.  
_This is the data for the poverty related analysis._  
```{r ,echo=FALSE}
# Assign variable with designated country names.
country01 <- c("United States", "Rwanda", "Mongolia", "Pakistan", "Lao PDR", "Bhutan", "Malaysia", "Brazil", "Ireland", "Japan", "Sweden", "Netherlands","Korea, Rep.")
# Assign variable with IDs.
wdi_ID_pov <- c("NY.GDP.MKTP.CD","NY.GDP.PCAP.CD","SP.POP.TOTL","SP.DYN.IMRT.IN","SP.DYN.LE00.IN","SH.STA.MMRT.NE")
# Assign variable to hold the data that I intend to analyze.
wdi_dat_pov <- WDI(indicator = wdi_ID_pov, start = 1960, end = 2016, extra = TRUE)
# remove all variables that are aggregated regional values.
wdi_dat_pov <- subset(wdi_dat_pov, region != "Aggregates")
# Rename the indicators to something simple.
names(wdi_dat_pov)[which(names(wdi_dat_pov) == "NY.GDP.MKTP.CD")] <- "GDP"
names(wdi_dat_pov)[which(names(wdi_dat_pov) == "NY.GDP.PCAP.CD")] <- "GDP_perCap"
names(wdi_dat_pov)[which(names(wdi_dat_pov) == "SP.POP.TOTL")] <- "Population"
names(wdi_dat_pov)[which(names(wdi_dat_pov) == "SP.DYN.LE00.IN")] <- "life_expectancy" 
names(wdi_dat_pov)[which(names(wdi_dat_pov) == "SP.DYN.IMRT.IN")] <- "infant_mortality"
names(wdi_dat_pov)[which(names(wdi_dat_pov) == "SH.STA.MMRT.NE")] <- "maternal_mortality"
```

```{r ,echo=FALSE}
##wdi_pov <- WDI(country="all", indicator = c("NY.GDP.PCAP.CD", "SP.DYN.LE00.IN", "SP.DYN.IMRT.IN"), start = 1960, end = 2017, extra = TRUE)
##wdi_dat_test <- WDI(country=c("US","KR"), indicator= c("SP.POP.TOTL", "SP.URB.TOTL.IN.ZS", "SP.RUR.TOTL.ZS"), start=1960, end=2017,extra=TRUE, cache=NULL)
##names(wdi_dat)
```

# Plot a graph

## Infant mortality
```{r ,echo=FALSE}
ggplot(subset(wdi_dat_pov, year == 2008), aes(x = GDP_perCap, y = infant_mortality)) + geom_point() +
  labs(x="GDP per Capita", y="Infant Mortality Rate", 
       title="Relationship between Infant Mortality Rate and GDP per Capita in year 2008")
```

## Life expectancy
```{r ,echo=FALSE}
co <- c("United States", "Rwanda", "Mongolia", "Pakistan", "Lao PDR", "Bhutan", "Malaysia", "Brazil", "Ireland", "Japan", "China","Sweden", "Netherlands","Korea, Rep.")
lifexp_c <- subset(wdi_dat_pov, country %in% co) # subset specific countries from the data
lifexp_c0 <- subset(lifexp_c, year == 2016)

ggplot(lifexp_c0) +
 aes(x = GDP_perCap, y = life_expectancy, colour = country) +
 geom_point(size = 2.56) +
 geom_text(aes(label = country), size=2, nudge_y = 0.8) +
 scale_color_hue() +
 labs(x = "GDP per Capita (USD)", y = "Life Expectancy (year)", title = "Figure 1: Life Expectancy and GDP per Capita by Country", subtitle = "(2016)", color = "Country") +
 scale_x_continuous(limits = c(0, 70000)) +
 theme_classic()
```

## Infant Mortality
```{r ,echo=FALSE}
ggplot(lifexp_c0) +
 aes(x = GDP_perCap, y = infant_mortality, colour = country) +
 geom_point(size = 2.56) +
 geom_text(aes(label = country), size=2, nudge_y = 1.9) +
 scale_color_hue() +
 labs(x = "GDP per Capita (USD)", y = "Infant Mortality Rate per 1,000 births", title = "Figure 2: Infant Mortality Rates by GDP per Capita", subtitle = "(2016)", color = "Country") +
 scale_x_continuous(limits = c(0, 70000)) +
 theme_classic()
```

## Simple Interactivity
```{r ,echo=FALSE}
library(plotly)
g <- ggplot(subset(lifexp_c, year > 1999 & year <= 2015), aes(x = GDP_perCap, y = infant_mortality, color = country == "Korea, Rep.", tooltip = country)) +
  geom_point(size = 0.8) +
  facet_wrap(~ year) + theme(legend.position="none") +
  labs(x="GDP per Capita (USD)", y="Infant Mortality Rate per 1,000 births", title= "Figure 3: GDP per capita vs Infant Mortality Rates by years",
       subtitle = "The colored dot represents Korea, Rep.", color = "Country") +
  scale_color_manual(labels = c("T999", "T888"), values = c("darkblue", "red")) +
  theme_dark()
ggplotly(g)
```

